Functory: A Distributed Computing Library \\ for Objective Caml \thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2011-02-25 doc/tfp-2011/main.tex
*** empty log message ***

@@

Our library, \functory, is implemented in \Ocaml\ and is available
from \url{http://www.lri.fr/~filliatr/functory/}.
In this section, we show the experimental results pertaining to the
case studies of Section~\ref{sec:studies}.

% \section{Illustrative Case Studies}\label{sec:studies}

In this section, we show how to use \functory\ on several case studies.
We focus here on the use of the API. Experimental results of these
case studies are given in Section~\ref{sec:experiments}.
The source code for all these case studies is contained in the
distribution, in sub-directory \texttt{tests}.

As a first example, let us consider matrix multiplication.
We assume two matrices \of{a} and \of{b}, respectively of size
$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
the standard matrix multiplication would be as follows:
\begin{ocaml}
  for i = 0 to n-1 do
    for j = 0 to m-1 do
      for k = 0 to p-1 do
	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
      done
    done
  done  
\end{ocaml}
and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
assuming addition and multiplication over coefficients to be 
denoted $+$ and $\times$ respectively.
An easy way to distribute this computation is obviously to turn each
computation of the inner loop over \of{k} into a task.
To do so, we first prepare the list of tasks:
\begin{ocaml}
  let tasks = 
    let l = ref [] in
    for i = 0 to n-1 do for j = 0 to m-1 do
      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
    done done;
    !l
\end{ocaml}
Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
tuple, together with position \of{(i,j)} as a second tuple.
The worker is receiving the first tuple as argument and computes
the dot product:
\begin{ocaml}
  let worker (ai, bj) =
    let c = ref 0 in
    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
    !c
\end{ocaml}
The master is a one line function which receives the result \of{r}
from the worker and simply stores it into \of{c}, according to the
position contained in the input task. It produces no new task.
\begin{ocaml}
  let master (_, (i,j)) r = c.(i).(j) <- r; []
\end{ocaml}
Finally, the overall computation is started by invoking \of{compute}
as follows:
\begin{ocaml}
  let () = compute ~worker ~master !tasks
\end{ocaml}
The total number of tasks is $\of{n}\times\of{m}$.

When master and worker programs are compiled with different
versions of the \Ocaml\ compiler, our library still provides
a monomorphic API over strings. As a consequence, we need to convert
tasks and results to and from strings in both master and worker.
The modified worker program then looks as follows:
\begin{ocaml}
  open Mono
  let worker (ai, bj) = ...
  let worker_string s = string_of_coeff (worker (task_of_string s))
  let () = Worker.compute worker_string ()
\end{ocaml}
The master program is modified in a similar way.
We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
as strings, as follows:
\begin{ocaml}
  let tasks = ... string_of_task ...
  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
\end{ocaml}
where \of{string_of_task} and \of{task_of_string} are user-defined
functions to convert tasks to and from strings.

@@

The next example is the classical $N$-queens problem, where we
compute the total number of ways to place $N$ queens on a $N\times N$
chessboard in such a way no two queens attack each other.
We use a standard backtracking algorithm for this problem, which
places the queens one by one starting from the first row.
Distributing the computation is thus quite easy: we consider all
possible ways to place queens on the first $D$ rows and then perform
the subsequent search in parallel. Choosing $D=1$ will result in
exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; and
so on.
% TODO: make a picture to illustrate positioning of 2 queens on the
% first 2 rows

@@

Drawing the Mandelbrot set is another classical example that could be
distributed easily, since the color of each point can be computed
independently of the others. 
Let us assume we are given the coordinates of the region to be drawn,
along with the width \of{w} and height \of{h} of the resulting image
(in pixels). 
Let us assume the total number of tasks $\of{t} \ge 1$ is given as a
parameter. It is immediate to split the image into \of{t} sub-images,
each of which is computed in parallel with and independently of the
others. For instance, the image could be split into horizontal slices
or, more generally, into rectangular blocks.

Each task is thus four floating-point numbers denoting the region
coordinates, together with two integers denoting the dimensions of the
sub-image to be drawn. The result of the task is a matrix of pixels,
of size $(\of{w}\times\of{h}) / \of{t}$. 
For instance, drawing a $800\times 600$ image using 20 tasks will
result in 20 sub-images of size $176,000$ bytes each,
assuming each pixel is encoded in four bytes.

In the network configuration, we deliberately choose to have two
different programs for master and workers, using the \of{Network.Poly}
module (actually, a single source code with a command line option).

% \subsection{Polynomial Multiplication}
% TODO?

\subsection{N-queens}

@@

This benchmark consists in drawing the fragment of the Mandelbrot set
with lower left corner $(-1.1, 0.2)$ and upper right corner $(-0.8,
0.4)$, as a $9,000\times6,000$ image. The sequential computation of
this image consumes 29.4 seconds. For \of{Cores} and \of{Network}
implementations, the computation times in seconds are tabulated below.

\subsection{Matrix Multiplication}

This benchmark was inspired by the PASCO'10 programming contest~\cite{PASCO}.
It consists of multiplication of
two square matrices of dimension 100, that is $\of{n} = \of{p} =
\of{m} = 100$, with integer coefficients.
Coefficients have several thousands of digits, hence we use
GMP~\cite{GMP} to handle operations over coefficients.

We compare the performances of two different implementations. In the
first one, called \of{mm1}, each task consists of the computation of a single
coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
In the second one, called \of{mm2}, each task consists of the
computation of a whole row of the resultant matrix.
As a consequence, the total number of tasks in
$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
The experimental results (in seconds) are tabulated below.
\begin{center}
  \begin{tabular}{|r|r|r|}
    \hline
    & \of{mm1}       & \of{mm2}  \\
    & (10,000 tasks) & (100 tasks) \\
    \hline\hline
\of{Sequential} & 20.3 &  20.2 \\
\hline
 \of{Cores} 
 (2 cores)     &   22.7 (0.89$\times$) &  11.3 (1.79$\times$) \\
 (4 cores)     &   12.3 (1.65$\times$) &   6.1 (3.31$\times$) \\
 (6 cores)     &    8.6 (2.36$\times$) &   4.3 (4.70$\times$) \\
 (8 cores)     &    8.0 (2.54$\times$) &   3.5 (\underline{5.77}$\times$) \\
 \hline
  \end{tabular}
\end{center}
We do not include results for the network configuration, as they do
not achieve any benefit with respect to the sequential
implementation. The reason is that the communication cost dominates the
computation cost in such a way that the total execution time is
always greater than 30 seconds. Indeed, irrespective of the
implementation (\of{mm1} or \of{mm2}), the total size of the
transmitted data is 
$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
billions of bytes.

A less naive implementation would have the worker read the input matrices
only once, \emph{e.g.} from a file, and then have the master send only
row and column indices. This would reduce the communication cost to
$O(\of{n}\times\of{m})$ only.

@@

>>>>

% TODO
% Our library, \functory, is implemented in \Ocaml\ and is available
% from \url{http://www.lri.fr/~filliatr/functory/}.

@@

In this section, we demonstrate the potential of using \functory\ on
several case studies.
The source code for all these case studies is contained in the
distribution, in sub-directory \texttt{tests/}.

@@

The first example is the classical $N$-queens problem, where we
compute the total number of ways to place $N$ queens on a $N\times N$
chessboard in such a way no two queens attack each other.
We use a standard backtracking algorithm for this problem, which
places the queens one by one starting from the first row.
Distributing the computation is thus quite easy: we consider all
possible ways to place queens on the first $D$ rows and then perform
the subsequent search in parallel. Choosing $D=1$ will result in
exactly $N$ tasks; choosing $D=2$ will result in $N^2-3N+2$ tasks; and
so on.
% TODO: make a picture to illustrate positioning of 2 queens on the
% first 2 rows

@@

@@

This benchmark was inspired by the PASCO'10 programming contest~\cite{PASCO}.
It consists of multiplication of
two square matrices of dimension 100 with integer coefficients.
Coefficients have several thousands of digits, hence we use
GMP~\cite{GMP} to handle operations over coefficients.

We compare the performances of two different implementations. In the
first one, called \of{mm1}, each task consists of the computation of a single
coefficient of the resultant matrix. 
In the second one, called \of{mm2}, each task consists of the
computation of a whole row of the resultant matrix.
As a consequence, the total number of tasks in
$10,000$ for \of{mm1} and only $100$ for \of{mm2}.
On the contrary, each task result for \of{mm1} is a single integer,
while for \of{mm2} it is a row of 100 integers.
%
The experimental results (in seconds) are tabulated below.
\begin{center}
  \begin{tabular}{|r|r|r|}
    \hline
    & \of{mm1}       & \of{mm2}  \\
    & (10,000 tasks) & (100 tasks) \\
    \hline\hline
\of{Sequential} & 20.3 &  20.2 \\
\hline
 \of{Cores} 
 (2 cores)     &   22.7 (0.89$\times$) &  11.3 (1.79$\times$) \\
 (4 cores)     &   12.3 (1.65$\times$) &   6.1 (3.31$\times$) \\
 (6 cores)     &    8.6 (2.36$\times$) &   4.3 (4.70$\times$) \\
 (8 cores)     &    8.0 (2.54$\times$) &   3.5 (\underline{5.77}$\times$) \\
 \hline
  \end{tabular}
\end{center}
We do not include results for the network configuration, as they do
not achieve any benefit with respect to the sequential
implementation. The reason is that the communication cost dominates the
computation cost in such a way that the total execution time is
always greater than 30 seconds. Indeed, irrespective of the
implementation (\of{mm1} or \of{mm2}), the total size of the
transmitted data is $10^6$ integers, which in our case amounts to
billions of bytes.

A less naive implementation would have the worker read the input matrices
only once, \emph{e.g.} from a file, and then have the master send only
row and column indices. This would reduce the amount of transmitted data to
$10,000$ integers only.

@@

\begin{comment}
  As a first example, let us consider matrix multiplication.  We
  assume two matrices \of{a} and \of{b}, respectively of size
  $\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as
  well as a matrix \of{c} of size $\of{n}\times\of{m}$ to store the
  result.  Assuming \of{a} being a row-matrix and \of{b} a
  column-matrix, the standard matrix multiplication would be as
  follows:
  \begin{ocaml}
    for i = 0 to n-1 do for j = 0 to m-1 do for k = 0 to p-1 do
    c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k) done done done
  \end{ocaml}
  and runs in $O(\of{n}\times\of{p}\times\of{m})$, assuming addition
  and multiplication over coefficients to be denoted $+$ and $\times$
  respectively.  An easy way to distribute this computation is
  obviously to turn each computation of the inner loop over \of{k}
  into a task.  To do so, we first prepare the list of tasks:
  \begin{ocaml}
    let tasks = let l = ref [] in for i = 0 to n-1 do for j = 0 to m-1
    do tasks := ((a.(i), b.(j)), (i,j)) :: !tasks done done; !l
  \end{ocaml}
  Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
  tuple, together with position \of{(i,j)} as a second tuple.  The
  worker is receiving the first tuple as argument and computes the dot
  product:
  \begin{ocaml}
    let worker (ai, bj) = let c = ref 0 in for k = 0 to p-1 do c := !c
    + ai.(k) * bj.(k) done; !c
  \end{ocaml}
  The master is a one line function which receives the result \of{r}
  from the worker and simply stores it into \of{c}, according to the
  position contained in the input task. It produces no new task.
  \begin{ocaml}
    let master (_, (i,j)) r = c.(i).(j) <- r; []
  \end{ocaml}
  Finally, the overall computation is started by invoking \of{compute}
  as follows:
  \begin{ocaml}
    let () = compute ~worker ~master !tasks
  \end{ocaml}
  The total number of tasks is $\of{n}\times\of{m}$.

  When master and worker programs are compiled with different versions
  of the \Ocaml\ compiler, our library still provides a monomorphic
  API over strings. As a consequence, we need to convert tasks and
  results to and from strings in both master and worker.  The modified
  worker program then looks as follows:
  \begin{ocaml}
    open Mono let worker (ai, bj) = ...  let worker_string s =
    string_of_coeff (worker (task_of_string s)) let () =
    Worker.compute worker_string ()
  \end{ocaml}
  The master program is modified in a similar way.  We simply replace
  \of{Poly} with \of{Mono} and encode/decode coefficients as strings,
  as follows:
  \begin{ocaml}
    let tasks = ... string_of_task ...  let master (_, (i,j)) r =
    c.(i).(j) <- coeff_of_string r; []
  \end{ocaml}
  where \of{string_of_task} and \of{task_of_string} are user-defined
  functions to convert tasks to and from strings.
\end{comment}

Drawing the Mandelbrot set is another classical example that could be
distributed easily, since the color of each point can be computed
independently of the others. 
This benchmark consists in drawing the fragment of the Mandelbrot set
with lower left corner $(-1.1, 0.2)$ and upper right corner $(-0.8,
0.4)$, as a $9,000\times6,000$ image. 
%
If the total number of tasks $\of{t} \ge 1$ is given as a
parameter,  it is immediate to split the image into \of{t} sub-images,
each of which is computed in parallel with and independently of the
others. In our case, the image is split into horizontal slices.
Each task is thus four floating-point numbers denoting the region
coordinates, together with two integers denoting the dimensions of the
sub-image to be drawn. The result of the task is a matrix of pixels,
of size $54,000,000 / \of{t}$. 
For instance, using $\of{t}=20$ tasks will
result in 20 sub-images of size $10.3$ Mb each,
assuming each pixel is encoded in four bytes.

The sequential computation of
this image consumes 29.4 seconds. For \of{Cores} and \of{Network}
implementations, the computation times in seconds are tabulated below.

% ICI
